---
title: "Gov 50 Final Project"
author: "Julia Dan"
description: "Gov 50 Final Project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

For my final project in Gov 50, I'd like to investigate data relating to the rate at which immigrants within the United States move back to their home country. I'd like to see whether there is a trend among immigrants from a certain world region returning to their country of origin as well as the ages at which an individual is more likely to return. Ultimately, I am interested in this field of inquiry because of my own familial background and trends I see among people of my community. I hope to run into incongruences in my data, and figure out how/why these come about. I hope to gather my data from the Migration Policy Institute (and various similar websites), and narrow down my data by year and region of the world (I might focus on Europe and other regions of high immigration). Through my analysis, I hope to at least see spikes in the volume of immigration to the United States by year in order to support my hypothesis that, while there is an overall increase in immigration since X year, there is also an increase of ex-patriots returning home. 


Milestone 2: Finding data and writing a proposal (due 10/13)
Proposal:
To what extent is immigration judge ideology correlated to the rate at which asylum is granted? In this study, I plan to answer this question, examining the relationship between different immigration judges’ characteristics (namely ideology and the party to which Congress belongs) influences the influx of immigrants into the United States. I hypothesize that a more democratic congress and liberal ideology score will result in higher immigration. I also plan to take in gender as well as English-speaking/non-English speaking status as a possible influence, should I find little correlation solely on ideology. These other variables do indicate a certain bias that could be connected with judge implicit ideology and decision-making. My sample is comprised of continuous counters of months elapsed starting with January 1990 as 1 and running through December 2010 as 252 based on the date the applicant’s case was decided. My explanatory variable of interest is ideology, and is based on the individual-level asylum_liberalism score; which ranges from 0.95 to 1.39 where higher scores indicating a greater proclivity to grant an applicant asylum. My outcome variable of interest is Ij_total_grants_relief: total grants of relief (including asylum and either type of withholding) by IJs in a given month. The observable pattern in my data that would prove my hypothesis is higher ideology scores correlate to higher total grants for relief. If this correlation is not seen, or even if a correlation between a Democratic Congress and more asylum acceptance is not seen, this would disprove my hypothesis completely.


Milestone 3:
A few things were re-worked in this iteration of my final project. For one, I am using a different data set that shows Supreme Court Justices and their ideology (according to Martin-Quinn judicial ideology scores) from 1937-2015. In this visualization, I tried to show the ideology of justices only from 2015. My data source can be found here: http://scdb.wustl.edu/documentation.php?var=issueArea and here: https://www.kaggle.com/datasets/umichigan/court-justices/.

```{r}
library(readxl)
scores <- read_excel("~/Desktop/scores.xlsx")
```

```{r}
library(ggplot2)
library(tidyverse)
```

```{r}
justices_2015 <- c("Anthony Kennedy", "Antonin Scalia", "Clarence Thomas", "Elena Kagan", "John Roberts", "Ruth Bader Ginsburg", "Samuel Alito", "Sonia Sotomayor", "Stephen Breyer")
```

```{r}
filtered_scores <- scores |>
  filter(scores == 2015)
```

```{r}
ggplot(data = filtered_scores, aes(x = ...2, y = ...4)) +
  geom_point() +
  labs(x = "Justice Code", y = "Mean Ideology") +
  ggtitle("Scatter Plot of Justice Name vs. Mean Ideology")
```



